\chapter{Security-constrained optimal power flow (\scopflow)}\label{chap:scopflow}
SCOPFLOW solves a contingency-constrained optimal power flow problem. The problem is set up as a two-stage optimization problem where the first-stage (base-case) represents the normal operation of the grid and the second-stage comprises $N_c$ contingency scenarios. Each contingency scenario can be single or multi-period.

\section{Formulation}

\subsection{Single-period}

The contingency-constrained optimal power flow (popularly termed as security-constrained optimal power flow (SCOPF) in power system parlance) attempts to find a least cost dispatch for the base case (or no contingency) while ensuring that if any of contingencies do occur then the system will be secure. This is illustrated in Fig. \ref{fig:scopflow} for a SCOPF with a base-case $c_0$ and three contingencies.

\input{figures/scopflow_fig.tex}

In general form, the equations for contingency-constrained optimal power flow are given by
(\ref{eq:scopflow_start}) -- (\ref{eq:scopflow_end}). This is a two-stage
stochastic optimization problem where the first stage is the base case $c_0$ and
each of the contingency states $c_i, i \in [1,N_c]$ are second-stage
subproblems. SCOPFLOW aims to minimize the objective $\sum_{c=0}^{N_c}f(x_c)$,
while adhering to the equality $g_c(x_c)$, inequality $h_c(x_c)$, and the
lower/upper bound ($x^-$,$x^+$) constraints. Equation (\ref{eq:scopflow_end})
represents the coupling between the base-case $c_0$ and each of the contingency states
$c_i$. Equation (\ref{eq:scopflow_end}) is the most typical form of coupling
that limits the deviation of the contingency variables $x_c$ from the base $x_0$
to within $\Delta x_c$. An example of this constraint could be the allowed real power output deviation for the generators constrained by their ramp limit, which is currently the only constraint SCOPFLOW supports.


\begin{align}
\centering
\text{min}&~\sum_{c=0}^{N_c}f_c(x_c)&  \label{eq:scopflow_start}\\
&\text{s.t.}& \nonumber \\
&~g_c(x_c) = 0,                             &c \in \left[0,N_c\right]& \\
&~h_c(x_c) \le 0,                           &c \in \left[0,N_c\right]& \\
x^- & \le x_c \le x^+,                     &c\in \left[0,N_c\right]& \\
-\Delta x_c & \le x_c - x_0 \le \Delta x_c,&c \in \left[1,N_c\right]&
\label{eq:scopflow_end}
\end{align}

\subsection{Multiperiod}

In the multi-period version, each contingency is comprised of multiple time-periods. The multiple periods have variables and constraints as described in chapter \ref{chap:tcopflow}. An example of multi-contingency multi-period optimal power flow is illustrated in Fig. \ref{fig:ctopflow} for multi-period SCOPFLOW with three contingencies $c_1$, $c_2$, and $c_3$ coupled to the base case $c_0$. Each state is multi-period with two time-periods. Each time-step is coupled with its adjacent one through ramping constraints. We assume that the contingency is incident at the first time-step, i.e. at $t_0$. This results in the coupling between the contingency cases $c_i, i \in [1,N_c]$ and the base-case $c_0$ only at time-step $t_0$.

\input{figures/ctopflow_fig}

The overall objective of this contingency-constrained multi-period optimal power flow is to find a secure dispatch for base-case $c_0$ while adhering to contingency and temporal constraints. The general formulation of this problem is given in Eqs. (\ref{eq:ctopflow_start}) -- (\ref{eq:ctopflow_end}).

\begin{align}
\centering
\text{min}&~\sum_{c=0}^{N_c}\sum_{t=0}^{N_t-1}f_{ct}(x_{c,t})& \label{eq:ctopflow_start}\\
&\text{s.t.}& \nonumber \\
&~g_{ct}(x_{c,t}) = 0,                                        &c \in \left[0,N_c\right], t \in \left[0,N_t-1\right]& \\
&~h_{ct}(x_{c,t}) \le 0,                                      &c \in \left[0,N_c\right], t \in \left[0,N_t-1\right]& \\
x^- & \le x_{c,t} \le x^+,                               &c \in \left[0,N_c\right], t\in \left[0,N_t-1\right]& \\
-\Delta x_t & \le x_{c,t} - x_{c,t-\Delta{t}} \le \Delta x_t,&c \in \left[0,N_c\right], t \in \left[1,N_t-1\right]& \label{eq:ctopflow_time_coupling}\\
-\Delta x_c & \le x_{c,0} - x_{0,0} \le \Delta x_c,&c \in \left[1,N_c\right]& \\
\label{eq:ctopflow_end}
\end{align}

In this formulation, the objective is to reduce the cost for the base-case time
horizon, where $f(x_{0,t})$ is the objective cost for contingency $c_0$ at time
$t$. Equation (\ref{eq:ctopflow_end}) represents the coupling between the base
case $c_0$ and each contingency $c_i$ at time-step $t_0$. We use a simple box
constraint $\Delta x_c$ to restrict the  deviation of decision variables
$x_{c,0}$ from the base-case $x_{0,0}$. The bound $\Delta x_c$ could represent here, for example, the allowable reserve for each generator.

\section{Solvers}
\scopflow supports solving the optimization problem via \ipopt, \hiop, or \emph{EMPAR}. \ipopt can solve \scopflow on single rank only. \hiop supports solving the problem in parallel using a primal-decomposition algorithm. With HIOP, one can solve the subproblem either on the CPU or GPU by selecting the appropriate subproblem model and solver (see options table below).  However, note that \emph{ExaGO needs to be built with \ipopt even when using \hiop solver.} The \emph{EMPAR} solver does not solve the security-constrained ACOPF problem, it only solves the base-case and the contingencies independently with \opflow. It distributes the contingencies to different processes when executed in parallel.

\section{Input and Output}
To execute SCOPFLOW, the following files are required:
\begin{itemize}
    \item \textbf{Network file:} The network file describing the network details. Only \matpower format files are currently supported.
    \item \textbf{Contingency file:} The file describing the contingencies. Contingencies can be single or multiple outages. The contingency file needs to be described in PTI format.
\end{itemize}
If the multi-period option is chosen, then additional files describing the load and wind generation can be (optionally) set.
\begin{itemize}
    \item \textbf{Load data:} One file for load real power and one for reactive power. The files need to be in CSV format. An example of the format for the 9-bus case is \href{https://gitlab.pnnl.gov/exasgd/frameworks/exago/-/tree/master/datafiles/case9}{here}.
    \item \textbf{Wind generation:} The wind generation time-series described in CSV format. See an example of the format \href{https://gitlab.pnnl.gov/exasgd/frameworks/exago/-/tree/master/datafiles/case9}{here}.
\end{itemize}

The \scopflow output is saved to a directory named \texttt{scopflowout}. This
directory contains $N_c$ files to save the solution for each contingency in
MATPOWER datafile format. Each file has the name \texttt{cont_xx} where \texttt{xx} is the contingency number. 

If the multi-period option is chosen then $N_c$ subdirectories are created (one
for each contingency), and each subdirectory contains $N_t$ output files, one
for each time-period. The subdirectories have the naming convention
\texttt{cont_xx} and the output file are named as \texttt{t_yy} where \texttt{yy} is the time-step number.

\section{Usage}

\begin{lstlisting}
  ./scopflow -netfile <netfilename> -ctgcfile <ctgcfilename> \
  <scopflowoptions> [-scopflow_enable_multiperiod 1]
\end{lstlisting}

\section{Options}
See table \ref{tab:scopflow_options}. In addition, all \opflow options in Table \ref{tab:opflow_options} and \tcopflow options in Table \ref{tab:tcopflow_options} can be used.
\begin{table}[H]
  \caption{SCOPFLOW options}
  \small
  \begin{tabular}{|p{0.3\textwidth}|p{0.2\textwidth}|p{0.3\textwidth}|p{0.2\textwidth}|}
    \hline
    \textbf{Option} & \textbf{Meaning} & \textbf{Values (Default value)} & \textbf{Compatibility} \\ \hline
    -netfile & Network file name & string $<$ 4096 characters (\href{https://gitlab.pnnl.gov/exasgd/frameworks/exago/-/blob/master/datafiles/case9/case9mod_gen3_wind.m}{case9mod\_gen3\_wind.m}) &\\ \hline
    -ctgcfile & Contingency file name & string $<$ 4096 characters (\href{https://gitlab.pnnl.gov/exasgd/frameworks/exago/-/blob/master/datafiles/case9/case9.cont}{case9.cont}) &\\ \hline
    -print\_output & Print output to screen & 0 or 1 (0) &\\ \hline
    -save\_output & Save output to directory & 0 or 1 (0) & Format determined by OPFLOW option. \\ \hline
    -scopflow\_output\_directory & Output directory path & ``scopflowout'' & \\ \hline
    -scopflow\_model & SCOPFLOW model type & GENRAMP, GENRAMPT (GENRAMP) & \\ \hline
    -scopflow\_solver & Set solver for scopflow & IPOPT, HIOP, or EMPAR (IPOPT) &\\ \hline
    -scopflow\_subproblem\_solver & Set solver for subproblem & IPOPT or HIOP (IPOPT) &Only when using HIOP solver for SCOPFLOW \\ \hline
    -scopflow\_subproblem\_model & Set model for subproblem & See OPFLOW chapter &Only when using HIOP solver for SCOPFLOW \\ \hline
    -scopflow\_Nc & Number of contingencies & int (0. Passing -1 results in all contingencies in the file used) &\\ \hline
    -scopflow\_mode & Operation mode: Preventive or corrective & 0 or 1 (0) &\\ \hline
    -scopflow\_enable\_multiperiod & Multi-period SCOPFLOW & 0 or 1 (0) & IPOPT solver only\\ \hline
    -scopflow\_ploadprofile & Real power load profile & string (\href{https://gitlab.pnnl.gov/exasgd/frameworks/exago/-/blob/master/datafiles/case9/load_P.csv}{load\_P.csv}) &\\ \hline
    -scopflow\_qloadprofile & Reactive power load profile & string (\href{https://gitlab.pnnl.gov/exasgd/frameworks/exago/-/blob/master/datafiles/case9/load_Q.csv}{load\_Q.csv}) &\\ \hline
    -scopflow\_windgenprofile & Wind generation profile & string (\href{https://gitlab.pnnl.gov/exasgd/frameworks/exago/-/blob/master/datafiles/case9/scenarios_9bus.csv}{case9/scenarios\_9bus.csv}) &\\ \hline
    -scopflow\_dT & Length of time-step (minutes) & double (5.0) &\\ \hline
    -scopflow\_duration & Total duration (hours) & double (0.5) &\\ \hline 
    -scopflow\_tolerance & Optimization solver tolerance & double (1e-6) &\\ \hline 
  \end{tabular}
  \label{tab:scopflow_options}
\end{table}

Depending on the value chosen for \texttt{-scopflow_mode}, SCOPFLOW can operate
either in \emph{preventive} (mode = 0) or \emph{corrective} (mode = 1) mode. In
the preventive mode, the PV and PQ generator real power is fixed to its
corresponding base-case values. The generators at the reference bus pick up any
make-up power required for the contingency. The corrective mode allows deviation
of the PV and PQ generator real power from the base-case dispatch constrained by
its 30-min. ramp rate capability. The optimization decides the optimal
redispatch. One can have AGC control instead of having the generators proportionally share the deficit/excess power by using the option \emph(-opflow_use_agc).

The option \texttt{-scopflow_enable_multiperiod} 1 must be used in order to enable any of the options listed in table \ref{tab:scopflow_options} for multiperiod analysis.

\section{Examples}

Some \scopflow example runs are provided with some sample output. Options are the default options given in table \ref{tab:opflow_options}, \ref{tab:tcopflow_options} and \ref{tab:scopflow_options} unless otherwise specified. Sample output is generated by running examples in the installation directory.

Example using the \ipopt solver:

\input{output/scopflow_ipopt_output.tex}

Example using HIOP solver with IPOPT subproblem solver.
\input{output/scopflow_hiop_output.tex}
