\chapter{Multi-period optimal power flow (TCOPFLOW)}\label{chap:tcopflow}

TCOPFLOW solves a full AC multi-period optimal power flow problem with the objective of minimizing the total cost over the given time horizon while adhering to constraints for each period and between consecutive time-periods (ramping constraints). 

\section{Formulation}
The multi-period optimal power flow problem is a series of optimal power flow problems coupled via temporal constraints. The generator real power deviation ($p_{jt}^{\text{g}} - p_{jt-\Delta{t}}^{\text{g}}$) constrained within the ramp limits form the temporal constraints. An illustration of the temporal constraints is shown in Fig. \ref{fig:tcopflow} with four time steps. Each time-step $t$ is coupled with its preceding time $t-\Delta{t}$, where $\Delta{t}$ is the time-step where the objective is to find a least cost dispatch for the given time horizon.

\input{figures/tcopflow_fig}

In general form, the equations for multi-period optimal power flow are given by
(\ref{eq:tcopflow_start}) -- (\ref{eq:tcopflow_end}). TCOPFLOW solves to
minimize the total generation cost $\sum_{t=0}^{N_t-1}f(x_t)$ over the time
horizon, where $N_t$ is the number of time-steps. At each time-step, the
equality constraints ($g(x_t)$), inequality $h(x_t)$, and the lower/upper limit
($x^-$,$x^+$) constraints need to be satisfied. Equation (\ref{eq:tcopflow_end})
represents the coupling between the consecutive time-steps. It is the most
common form of coupling that limits the deviation of the real power generation
at time $t$ from its preceding time-step $t-\Delta{t}$ to within its ramping
capability $\Delta x_t$.


\begin{align}
\centering
\text{min}&~\sum_{t=0}^{N_t-1} f(x_t) &  \label{eq:tcopflow_start}\\
&\text{s.t.}& \nonumber \\
&~g(x_t) = 0,                                        &t \in \left[0,N_t-1\right]& \\
&~h(x_t) \le 0,                                      &t \in \left[0,N_t-1\right]& \\
x^- & \le x_t \le x^+,                               &t\in \left[0,N_t-1\right]& \\
-\Delta x_t & \le x_t - x_{t-\Delta{t}} \le \Delta x_t,&t \in \left[1,N_t-1\right]&
\label{eq:tcopflow_end}
\end{align}

\section{Solvers}\label{sec:tcopflow_solvers}%
Currently, \exago only supports solving \tcopflow using \ipopt~ on on a single rank. %Option:\\ %\opflowoption{-tcopflow_solver IPOPT}

\section{Input and Output}
\begin{itemize}
    \item \textbf{Network file:} The network file describing the network details. Only \matpower format files are currently supported.
    \item \textbf{Load data:} One file for load real power and one for reactive power. The files need to be in CSV format. An example of the format for the 9-bus case is \href{https://gitlab.pnnl.gov/exasgd/frameworks/exago/-/tree/master/datafiles/case9}{here}.
    \item \textbf{Wind generation:} The wind generation time-series described in CSV format. See an example of the format \href{https://gitlab.pnnl.gov/exasgd/frameworks/exago/-/tree/master/datafiles/case9}{here}.
\end{itemize}
If the load data and/or wind generation profiles are not set then a flat profile is assumed, i.e., the load and wind generation for all hours is constant.

The \tcopflow output is saved to a directory named \texttt{tcopflowout}. This directory contains $N_t$ files, one for each time-step, in \matpower data file format.

\section{Usage}
\begin{lstlisting}
\input{output/tcopflow_output.tex}
\end{lstlisting}

